---
title: "Formatting options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(clerkR)
```

## Overview

All `tbl_*` constructors inherit their formatting defaults from
`clerk_options()`. Defaults follow biomed/APA conventions and are loaded
automatically when the package is attached — no setup required. You can
inspect the current settings at any time:

```{r inspect}
clerk_options()
```

Options can be changed session-wide with a single call, or overridden
per-table by passing the same arguments directly to any `tbl_*` function.

---

## P-value style

The `p_style` option controls how p-values are displayed. Four styles are
available:

| Style | Example output | Use case |
|---|---|---|
| `"apa"` | `= 0.032`, `< 0.001` | APA manuscripts (default) |
| `"plain"` | `0.032`, `< 0.001` | Plain decimal |
| `"stars"` | `**` | Stars only, no numeric p |
| `"stars_p"` | `= 0.032 **` | Stars alongside numeric p |

```{r p-style}
# APA style (default)
tbl_descriptive(clerk_example, group = sex, output = "gt") |>
  clerk_render(title = "APA style (default)")

# Plain decimal
tbl_descriptive(clerk_example, group = sex, output = "gt",
                p_style = "plain") |>
  clerk_render(title = "Plain decimal")
```

To set a different default for the whole session:

```{r set-style, eval = FALSE}
clerk_options(p_style = "plain")
```

---

## Decimal places

Three separate options control decimal places for different value types:

| Option | Default | Controls |
|---|---|---|
| `digits` | `2` | Continuous statistics (mean, SD, β, SE) |
| `r_digits` | `3` | Correlations and h² estimates |
| `p_digits` | `3` | P-values |

```{r digits}
# More decimal places for p-values, fewer for statistics
tbl_descriptive(clerk_example, group = sex, output = "gt",
                digits = 1, p_digits = 2) |>
  clerk_render(title = "digits = 1, p_digits = 2")
```

The `p_threshold` option controls the value below which p-values are shown
as `< {threshold}` rather than as a decimal. The default is `0.001`:

```{r threshold}
# Show < 0.01 instead of < 0.001
tbl_descriptive(clerk_example, group = sex, output = "gt",
                p_threshold = 0.01) |>
  clerk_render(title = "p_threshold = 0.01")
```

---

## Significance stars

Stars are off by default. Enable them with `stars = TRUE`:

```{r stars}
tbl_correlation(clerk_cor_example, fdr = TRUE, output = "gt",
                stars = TRUE) |>
  clerk_render(title = "With significance stars")
```

The thresholds for `*`, `**`, and `***` are set by `star_thresholds`
(default `c(0.05, 0.01, 0.001)`). For stars-only output use
`p_style = "stars"`, or combine stars with numeric p-values using
`p_style = "stars_p"`:

```{r stars-p}
tbl_regression(clerk_reg_example, output = "gt",
               p_style = "stars_p") |>
  clerk_render(title = "Stars alongside p-values")
```

---

## FDR display

When `fdr = TRUE` is passed to a `tbl_*` function, a `p (FDR)` column is
added containing BH-adjusted p-values. Three options control how non-surviving
results are displayed:

| Option | Default | Effect |
|---|---|---|
| `fdr_ns` | `TRUE` | Replace non-surviving cells with `fdr_ns_label` |
| `fdr_alpha` | `0.05` | Alpha level applied to the **BH-adjusted** p-value |
| `fdr_ns_label` | `"ns"` | The label shown for non-surviving cells |

```{r fdr-ns}
# Default: non-survivors labelled "ns"
tbl_descriptive(clerk_example, group = sex, fdr = TRUE, output = "gt") |>
  clerk_render(title = "FDR with ns label (default)")
```

Disable the replacement to always show the numeric FDR value:

```{r fdr-numeric}
tbl_descriptive(clerk_example, group = sex, fdr = TRUE, output = "gt",
                fdr_ns = FALSE) |>
  clerk_render(title = "FDR numeric values always shown")
```

Use a stricter survival threshold:

```{r fdr-strict}
tbl_correlation(clerk_cor_example, fdr = TRUE, output = "gt",
                fdr_alpha = 0.01) |>
  clerk_render(title = "FDR survival at \u03b1 = 0.01")
```

---

## Domain labels for unassigned variables

When a `domains` list is supplied, variables not assigned to any domain receive
a blank section header by default (`domain_other = ""`). Set it to a string
to collect them under a named section:

```{r domain-other}
tbl_descriptive(
  clerk_example,
  group        = sex,
  domains      = list("Metabolic" = c("hdl", "glucose", "bmi")),
  domain_other = "Other",
  output       = "gt"
) |>
  clerk_render(title = "Unassigned variables under 'Other'")
```

---

## Session-wide vs per-call options

Options set via `clerk_options()` apply to all subsequent tables in the
session. Arguments passed directly to a `tbl_*` function override the session
default for that call only:

```{r session-vs-call}
# Change session default to plain style
clerk_options(p_style = "plain")

# This table uses plain (from session)
tbl_simple(clerk_example, output = "gt") |>
  clerk_render(title = "Session default: plain")

# This table overrides back to APA for this call only
tbl_simple(clerk_example, output = "gt", p_style = "apa") |>
  clerk_render(title = "Per-call override: apa")

# Reset to factory defaults
clerk_options(reset = TRUE)
```

---

## Full option reference

```{r reference, eval = FALSE}
clerk_options(
  digits          = 2,        # continuous stats decimal places
  r_digits        = 3,        # correlations / h² decimal places
  p_digits        = 3,        # p-value decimal places
  p_threshold     = 0.001,    # below this → "< 0.001"
  p_style         = "apa",    # "apa" | "plain" | "stars" | "stars_p"
  stars           = FALSE,    # append * ** ***
  star_thresholds = c(0.05, 0.01, 0.001),
  fdr_ns          = TRUE,     # replace non-surviving FDR with label
  fdr_alpha       = 0.05,     # applied to BH-adjusted p, not raw p
  fdr_ns_label    = "ns",     # label for non-surviving FDR cells
  domain_other    = ""        # label for unassigned variables
)
```
