---
title: "Heritability tables with R-itable and clerkR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Heritability tables with R-itable and clerkR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

`clerkR::tbl_heritability()` is designed to accept the output of
`R-itable::herit_batch()` directly — no column renaming required. This
vignette shows the full workflow from pedigree to publication-ready table.

The pipeline has four steps:

1. Simulate (or load) a pedigree and phenotype data frame.
2. Build the additive genetic relationship matrix (GRM) with `build_grm()`.
3. Run batch heritability estimation with `herit_batch()`.
4. Pass the results straight into `tbl_heritability()` and `clerk_render()`.

## Setup

```{r libraries}
library(Ritable)
library(clerkR)
```

## 1. Simulate a family cohort

We simulate 80 nuclear families (2 parents + 3 offspring each) with three
quantitative traits at different true h² values — BMI (~0.60), systolic blood
pressure (~0.35), and HDL cholesterol (~0.50). This mirrors the structure of a
real family cohort study such as the Baependi Heart Study.

```{r simulate}
set.seed(2026)
n_fam <- 80

# Pedigree: founders have pat = 0 / mom = 0
ped <- do.call(rbind, lapply(seq_len(n_fam), function(f) {
  base <- (f - 1L) * 5L
  data.frame(
    id  = base + 1:5,
    pat = c(0L, 0L, base + 1L, base + 1L, base + 1L),
    mom = c(0L, 0L, base + 2L, base + 2L, base + 2L),
    sex = c(1L, 2L, sample(1:2, 3, replace = TRUE))
  )
}))

# Simulate trait values for all pedigree members
simulate_trait <- function(ped, n_fam, h2, seed) {
  set.seed(seed)
  genetic  <- rep(rnorm(n_fam, 0, sqrt(h2)),      each = 5)
  residual <- rnorm(nrow(ped),         0, sqrt(1 - h2))
  genetic + residual
}

pheno_all <- data.frame(
  bmi = simulate_trait(ped, n_fam, h2 = 0.60, seed = 1),
  sbp = simulate_trait(ped, n_fam, h2 = 0.35, seed = 2),
  hdl = simulate_trait(ped, n_fam, h2 = 0.50, seed = 3)
)

# Keep only offspring rows for the phenotype data frame
off_rows <- ped$pat != 0L
dat <- data.frame(
  IID     = ped$id[off_rows],
  age     = round(runif(sum(off_rows), 20, 75)),
  sex_num = ped$sex[off_rows],
  bmi     = pheno_all$bmi[off_rows],
  sbp     = pheno_all$sbp[off_rows],
  hdl     = pheno_all$hdl[off_rows]
)
dat$age2 <- dat$age^2
```

## 2. Build the GRM

`build_grm()` computes the additive genetic relationship matrix from the
pedigree. The full pedigree (founders + offspring) is passed so that shared
ancestry can be traced correctly; only the study subjects' IDs are returned in
the matrix.

```{r grm}
A <- build_grm(ped, study_ids = dat$IID)
```

## 3. Batch heritability estimation

`herit_batch()` iterates over all trait × covariate-model combinations and
returns a tidy data frame with one row per successfully fitted model.

```{r batch}
res <- herit_batch(
  traits    = c("bmi", "sbp", "hdl"),
  grm       = A,
  data      = dat,
  covs_list = list(
    unadj = NULL,
    cov1  = c("age", "sex_num"),
    cov2  = c("age", "sex_num", "age2")
  )
)

res
```

The output columns (`trait`, `covariates`, `h2`, `ci_lo`, `ci_hi`, `pval`,
`sigma2_a`, `sigma2_e`) match `tbl_heritability()` defaults exactly.

## 4. Render the table

Pipe the `herit_batch()` output directly into `tbl_heritability()`. The only
argument needed beyond the defaults is `model = covariates` to tell
`tbl_heritability()` which column identifies the covariate model.

```{r table}
res |>
  tbl_heritability(
    model    = covariates,
    sigma2_a = sigma2_a,
    sigma2_e = sigma2_e,
    domains  = list(
      "Cardiometabolic" = c("bmi", "sbp", "hdl")
    ),
    fdr      = TRUE,
    output   = "gt"
  ) |>
  clerk_render(
    title    = "Heritability estimates",
    subtitle = "Narrow-sense h\u00b2 \u00b1 95% profile-likelihood CI",
    footnote = "FDR correction applied within each covariate model (BH). \
p-values are one-sided LRT with chi-squared(1) boundary correction."
  )
```

## Filtering to a single model

When you want to show only one covariate model — for example the fully adjusted
model for a main results table — filter before passing to `tbl_heritability()`:

```{r filtered}
res |>
  subset(covariates == "cov2") |>
  tbl_heritability(
    sigma2_a = sigma2_a,
    sigma2_e = sigma2_e,
    fdr      = TRUE,
    output   = "gt"
  ) |>
  clerk_render(
    title    = "Heritability estimates (age + sex + age\u00b2)",
    footnote = "FDR correction applied across all traits (BH)."
  )
```

## HTML output

For an interactive table, set `output = "html"` at construction time:

```{r html}
res |>
  subset(covariates == "cov2") |>
  tbl_heritability(
    sigma2_a = sigma2_a,
    sigma2_e = sigma2_e,
    output   = "html"
  ) |>
  clerk_render()
```
