---
title: "Heritability tables with R-itable and clerkR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Heritability tables with R-itable and clerkR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
ritable_available <- requireNamespace("Ritable", quietly = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = ritable_available
)
```

```{r not-available, eval=!ritable_available, echo=FALSE, results="asis"}
cat("> **Note:** This vignette requires the `Ritable` package
> (`remotes::install_github(\"circadia-bio/R-itable\")`).
> Code is shown below but not executed in this build.")
```

## Overview

`clerkR::tbl_heritability()` is designed to accept the output of
`R-itable::herit_batch()` directly. Column-name defaults match `herit_batch()`
output exactly:

| `tbl_heritability()` argument | Default | `herit_batch()` column |
|---|---|---|
| `metric` | `"trait"` | trait name |
| `h2` | `"h2"` | heritability estimate |
| `ci_low` | `"ci_lo"` | lower CI bound |
| `ci_high` | `"ci_hi"` | upper CI bound |
| `p` | `"pval"` | one-sided LRT p-value |

The covariate model column is `"covariates"` in `herit_batch()` output — pass
`model = "covariates"` to include it.

## Setup

Install both packages from GitHub if you haven't already:

```{r install, eval=FALSE}
# install.packages("remotes")
remotes::install_github("circadia-bio/R-itable")
remotes::install_github("circadia-bio/clerkR")
```

```{r libraries}
library(Ritable)
library(clerkR)
```

## 1. Simulate a family cohort

We simulate 80 nuclear families (2 parents + 3 offspring each) with three
quantitative traits at different true h² values — BMI (~0.60), systolic blood
pressure (~0.35), and HDL cholesterol (~0.50).

```{r simulate}
set.seed(2026)
n_fam <- 80

ped <- do.call(rbind, lapply(seq_len(n_fam), function(f) {
  base <- (f - 1L) * 5L
  data.frame(
    id  = base + 1:5,
    pat = c(0L, 0L, base + 1L, base + 1L, base + 1L),
    mom = c(0L, 0L, base + 2L, base + 2L, base + 2L),
    sex = c(1L, 2L, sample(1:2, 3, replace = TRUE))
  )
}))

simulate_trait <- function(ped, n_fam, h2, seed) {
  set.seed(seed)
  genetic  <- rep(rnorm(n_fam, 0, sqrt(h2)),      each = 5)
  residual <- rnorm(nrow(ped),         0, sqrt(1 - h2))
  genetic + residual
}

pheno_all <- data.frame(
  bmi = simulate_trait(ped, n_fam, h2 = 0.60, seed = 1),
  sbp = simulate_trait(ped, n_fam, h2 = 0.35, seed = 2),
  hdl = simulate_trait(ped, n_fam, h2 = 0.50, seed = 3)
)

off_rows <- ped$pat != 0L
dat <- data.frame(
  IID     = ped$id[off_rows],
  age     = round(runif(sum(off_rows), 20, 75)),
  sex_num = ped$sex[off_rows],
  bmi     = pheno_all$bmi[off_rows],
  sbp     = pheno_all$sbp[off_rows],
  hdl     = pheno_all$hdl[off_rows]
)
dat$age2 <- dat$age^2
```

## 2. Build GRM and run batch estimation

```{r batch}
A <- build_grm(ped, study_ids = dat$IID)

res <- herit_batch(
  traits    = c("bmi", "sbp", "hdl"),
  grm       = A,
  data      = dat,
  covs_list = list(
    unadj = NULL,
    cov1  = c("age", "sex_num"),
    cov2  = c("age", "sex_num", "age2")
  )
)

res
```

## 3. Render the table

Pass column names as strings — defaults already match `herit_batch()` output,
so only `model`, `sigma2_a`, and `sigma2_e` need to be specified:

```{r table}
res |>
  tbl_heritability(
    model    = "covariates",
    sigma2_a = "sigma2_a",
    sigma2_e = "sigma2_e",
    domains  = list("Cardiometabolic" = c("bmi", "sbp", "hdl")),
    fdr      = TRUE,
    output   = "gt"
  ) |>
  clerk_render(
    title    = "Heritability estimates",
    subtitle = "Narrow-sense h\u00b2 \u00b1 95% profile-likelihood CI",
    footnote = "FDR correction applied within each covariate model (BH). p-values are one-sided LRT with chi-squared(1) boundary correction."
  )
```

## Filtering to a single model

```{r filtered}
res |>
  subset(covariates == "cov2") |>
  tbl_heritability(
    sigma2_a = "sigma2_a",
    sigma2_e = "sigma2_e",
    fdr      = TRUE,
    output   = "gt"
  ) |>
  clerk_render(
    title    = "Heritability estimates (age + sex + age\u00b2)",
    footnote = "FDR correction applied across all traits (BH)."
  )
```

## HTML output

```{r html}
res |>
  subset(covariates == "cov2") |>
  tbl_heritability(
    sigma2_a = "sigma2_a",
    sigma2_e = "sigma2_e",
    output   = "html"
  ) |>
  clerk_render()
```
